---
title: "Assignment 4"
author: "Akeem Shepherd"
date: "10/7/2021"
output: pdf_document
---
 A) Does age (medage16) appear to be associated with unemployment (pct_unemp)?
### B) Does race (pct_white) appear to be assocaited with educational attainment (pct_BAdeg)?
### C) Do these relationships appear to vary by population size (urban_class)?
### 
### To complete the HW 1) create an appropriate univariate plot to describe the distribution
### of all 5 of the variables in the analysis, 2) Create a bivariate plot describing the
### relationships of interest in A and B above, 3) Create a set of bivariate sub-plots for 
### A and B faceted by population size.
###
###SUbmit your R-Script in markdown form (.rmd) as your Lab/HW 4 Assignment Deliverable.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/OWNER-PC/Downloads")
getwd()
list.files()
require(data.table)
library(ggplot2)

mydata1 <- fread("Lab4a_data.csv", 
                 stringsAsFactors = F, 
                 data.table = F)

mydata2 <- fread("Lab4b_data.csv", 
                 stringsAsFactors = F, 
                 data.table = F)

names(mydata1)
names(mydata2)
head(mydata1)
head(mydata2)

mydata <- merge(mydata1,
                mydata2, 
                by = "id")

names(mydata)
head(mydata)
rm(mydata1,mydata2)

mydata <- mydata[,c(3,5,7,9,11,13,15,17,19)]

names(mydata)
head(mydata)

```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
####Clean and manage data
for(i in 2:8){
  print(i)
  pct_vars <- (mydata[i]/mydata$Totpop16)*100
  mydata <- cbind(mydata,pct_vars)
  pct_vars<-NULL
}
names(mydata)[10:15] <- c("pct_male", "pct_female", "pct_medage16", "pct_white","pct_black","pct_BAdeg","pct_unemp")
head(mydata)

```


```{r, EDA}
####Exploratory Data Analysis (EDA) and simple plots in R
####Univariate
summary(mydata)
summary(mydata$medinc16)

mydata$BA_binary <- ifelse(mydata$BAdeg16<mean(mydata$BAdeg16),
                           "Below Avg. % BA","Above Avg. % BA")
mydata$INC_binary <- ifelse(mydata$medinc16<mean(mydata$medinc16),
                            "Below Avg. Med Income","Above Avg Med Income")

table(mydata$BA_binary)
prop.table(table(mydata$BA_binary))
prop.table(table(mydata$BA_binary,mydata$INC_binary))

table <- table(mydata$BA_binary)

mydata$BA_binary_num <- ifelse(mydata$pct_BAdeg<mean(mydata$pct_BAdeg),
                           0,1)

hist(mydata$BA_binary_num)

hist(mydata$pct_white)

barplot(table)

barplot(table, 
        main="Frequency of Respondents Above/Below Avg. BA Degree Rate",
        ylab = "Frequency", 
        xlab = "Group")


```

```{r, EDA}
####Exploratory Data Analysis (EDA) and simple plots in R
####Univariate
summary(mydata$BAdeg16)
summary(mydata$medinc16)

mydata$BA_binary <- ifelse(mydata$BAdeg16<mean(mydata$BAdeg16),
                           "Below Avg. % BA","Above Avg. % BA")
mydata$INC_binary <- ifelse(mydata$medinc16<mean(mydata$medinc16),
                            "Below Avg. Med Income","Above Avg Med Income")

table(mydata$BA_binary)
prop.table(table(mydata$BA_binary))
prop.table(table(mydata$BA_binary,mydata$INC_binary))

table <- table(mydata$BA_binary)

mydata$BA_binary_num <- ifelse(mydata$pct_BAdeg<mean(mydata$pct_BAdeg),
                           0,1)

hist(mydata$BA_binary_num)

hist(mydata$pct_white)

barplot(table)

barplot(table, 
        main="Frequency of Respondents Above/Below Avg. BA Degree Rate",
        ylab = "Frequency", 
        xlab = "Group")
```
